Dark Control:\\ A Unified Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-05-03 paper/paper.tex
this and that

Normal and disturbed metabolic fluctuations in the
human PMC have been closely related to
changes of conscious awareness \citep{cavanna2006precuneus}.
Indeed,
the PMC matures relatively late (i.e., myelination) during postnatal development in monkeys
\citep{goldman1987development}, which is generally considered to
be a sign of evolutionary sophistication \citep{flechsig1920}.
%
This DMN node has long been speculated to
reflect constant computation of
environmental statistics and its internal representation
as an inner ``mind's eye'' \citep{cavanna2006precuneus, leech_pcc2014}.
For instance, B\'alint's syndrome is a neurological disorder of conscious
awareness that results from medial damage in the parietal cortex
\citep{balint1909seelenlahmung}.
Patients are plagued by an
inability to combine various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as an inability to direct action towards
currently unattended environmental objects
(i.e., optic ataxia).
This can be viewed as a high-level impairment in gathering
information about alternative objects (i.e., exploration) as well as
leveraging these environmental opportunities (i.e., exploitation).
Congruently,
the human PMC was coupled in two functional connectivity analyses
\citep{bzdok2015subspecialization}
with the amygdala, involved in significance evaluation, and
the nucleus accumbens (NAc), involved in reward evaluation.
Specifically, among all parts of the PMC,
the ventral posterior cingulate cortex was
most connected to the laterobasal
nuclei group of the amygdala
\citep{bzdok2015subspecialization}.
This amygdalar subregion has been proposed to
continuously scan environmental input
for biological significance assessment
\citep{trust_attract}.

@@

\section{Relation to existing cognitive accounts on default mode function}
\subsection{Predictive coding hypothesis}
Predictive coding mechanisms
\citep{clark2013whatever, friston2008hierarchical}
are a frequently evoked idea in the context of default mode function
\citep{bar2007}.
Cortical responses are explained as
emerging from continuous functional interaction between
higher and lower levels of the neural processing hierarchy.
Feed-forward sensory processing is constantly calibrated by
top-down modulation from more multi-sensory and associative brain regions
further away from primary sensory cortical regions.
The dynamic interplay between cortical processing levels
may enable inferences about the world by reconciling
gaps between fresh sensory input and expectations computed
based on stored prior information.
At each stage of neural processing,
an internally generated prediction of aspects of environmental sensations is
directly compared against the actual environmental input.
A prediction error at one of the processing levels
incurs plasticity changes of neuronal back projections
(i.e., adapting model parameters)
to allow for gradually improved future prediction of the environment.
In this way,
the predictive coding hypothesis offers explanations for
the constructive, generative nature of sensory perception
\citep{friston2010free} and
the intimate relation of motor action to sensory expectations
\citep{wolpert1995internal, kording2004bayesian}.
Contextual sensorimotor integration of perception-action cycles
may be maintained by top-down modulation
using a-priori information about the environment.

In short, predictive coding processes
conceptualize updates of the internal representation of the environment
to best accommodate and prepare the organism
for processing the constant influx of sensory stimulation and
performing action on the environment.
There are hence a number of common properties
between the predictive coding account
and the proposed formal account based on MDPs.
Importantly,
a generative model of how perceived sensory cues arise in the
world would be incorporated into
the current neuronal wiring.
Further,
both accounts are plausibilized by neuroscientific evidence that suggest
the human brain to be a ``statistical organ'' \citep{friston2014phantastic}
with the biological purpose to
generalize from the past to new experiences.
Neuroanatomically, axonal
back projections indeed outnumber by far the axonal input projections
existing in the monkey and probably also human brain
\citep{salin1995corticocortical}.
Moreover,
both accounts provide a parsimonious explanation why the
human brain decreases processing load devoted to incoming information
when the environment is predictable because the internal generative
model only requires updating following descrepancies between
environmental reality and its internal reinstantiation.
Increased computation resources are however allocated
when unknown stimuli or
unexpected events are encountered by the organism.
The predictive coding and formal account
naturally evoke a mechanism of brain plasticity in that
neuronal wiring gets increasingly adapted
when faced by unprecedented environmental challenges.

While sensory experience is a generative process from both views,
the predictive coding account frames
sensory perception of the external world
as a generative experience due to the modulatory top-down influence at
various stages of sensory input processing.
The sequential decision-making framework inherent in our MDP view of the DMN,
replaces such a top-down generative design.
Further,
the hierarchical processing aspect from predictive coding
is re-expressed in our account in form of
nested prediction of probable upcoming actions, states, and rewards.
Both accounts capture the consequences of action.
The predictive coding account is typically explained without
explicit parameterization of the agent's horizon and
has a tendency to be presented as emphasizing prediction about the
immediate future.
In the present account, the horizon of that
look into the future is made explicit in the $\gamma$ parameter
of the Bellman equation.
% \item \suggestremove{Mismatch negativity of PC is immanent in reward prediction error
% by the difference between actually predicted reward of an action given
% a state and the reward predicted by non-linear regression with graded
% discounting of future rewards.}
Finally,
the process of adapting the neuronal connections
for improved top-down modulation
takes the concrete form of gradient computation and
back-propagation in our MDP implementation.
It is however important to note that
the neurobiological plausibility of
the back-propagation procedure is controversal
\citep{goodfellow2016deep}.

In sum,
the present formal account may serve as a concrete implementation of
the long-standing predictive coding account from cognitive neuroscience.
MDPs have the advantage of exposing an explicit
mechanism for the horizon of future considerations or
how the internal model of the world gets updated,
and
more explicitly reconciles sensory input processing and action consideration.
%
Recasting DMN function in terms of MDPs therefore naturally incorporates
a majority of aspects from the prediction coding hypothesis.

\subsection{Semantic hypothesis}
Another frequently proposed cognitive account to explain DMN function revolves
around forming logical associations and abstract analogies between
experiences and conceptual knowledge derived from past behavior
\citep{bar2007proactive, binder1999conceptual, constantinescu2016organizing}.
Analogies might naturally tie incoming new sensory stimuli to
explicit world knowledge (i.e., semantics) about the environment
\citep{bar2009proactive}.
The encoding of complex environments could thus be facilitated
by association to similar states.
%
Going beyond meaning and concepts extracted from the world,
semantic building blocks may provide the basis for
mentally envisioning non-existing scenarios
that would optimize behavior
by constantly simulating possible future events
\citep{boyer2008evolutionary, binder2011neurobiology}.
Such cognitive processes can afford
the internal construction and elaboration of necessary information
that is not presented in the surrounding environment
by recombining building blocks of
concept knowledge and episodic memories
\citep{hassabis2009construction}.
Indeed, in aging humans, remembering the past and imaging the future
equally decreased in the level of detail and were associated with
concurrent deficits in forming and integrating relationships between
items \citep{addis2008age, spreng2006temporal}.
Further,
% a constructive account can explain the reciprocal relationship
% between an egocentric first person perspective and
% an allocentric bird’s eye perspective immersed in
% self-reflection, semantic associations, and autobiographical memories.
% %
% Cognitive aspects of egocentric-allocentric switching
episodic memory, language, problem solving,
planning, estimating other people's thoughts, and spatial navigation
represent neural processes that are likely to
build on abstract world knowledge and logical associations
for integrating the constituent elements in mental scene construction
\citep{schacter2007remembering}.
Such scene construction processes could contribute to interpreting the
present and foretelling the future.
Mental scene construction has been proposed
to imply a distinction between
engagement in the sensory environment
and internally generated mind-wandering
\citep{buckner2007self}.
These investigators stated that
``A computational model [...] will probably require a form of
regulation by which perception of the current world is suppressed
while simulation of possible alternatives are constructed,
followed by a return to perception of the present.''.
The semantic hypothesis is for instance supported by evidence in animals that
could learn a \textit{cognitive map of the environment}
and exploit it later for other means \citep{tolman1948cognitive}.

In comparison,
both the semantic hypothesis and the present formal account based on MDPs
expose mechanisms of how action considerations could be mentally explored.
In both accounts,
there is also no reason to assume that predictions of various
levels of complexity, abstraction, timescale and purpose
rely on mechanisms that are qualitatively different. This concurs with
DMN activity increases across time, space, and content domains
demonstrated in many neuroimaging studies
\citep{spreng2009common, laird2009, bzdok2012morality, binder2009}.
Further, the semantic hypothesis
and MDP account provide explanations why HC damage does
not only impair recalling memories, but also hypothetical and future
thinking \citep{hassabis2007patients}.
While both semantic hypothesis and
our formal account propose memory-based internally
generated information for probabilistic mental models of action outcomes,
MDPs render explicit the grounds on which the final action is
eventually chosen, namely, the estimated cumulative reward.
\textbf{The learned
estimations of value and action transitions drive
stimulus processing and action choice in the present.}
Additionally,
the MDPs naturally integrate the egocentric view
(more related to current action, state and reward) and the
world view (more related to past and future actions, states, and rewards)
on the world in a same optimization problem.
Finally,
the semantic account of DMN function does not offer
a mechanistic explanation \textit{how}
explicit world knowledge and semantic analogies thereof
lead to prediction of future actions and states.
The semantic hypothesis does not explain why memory recall
for scene construction in humans is typically fragmentary and noisy
instead of accurate and reliable.
In contrast to existing accounts on semantics and
mental scene construction, the random and creative aspects of DMN function
are explained in MDPs by the advantages of stochastic optimization.
Our MDP account provides an algorithmic explanation in that
stochasticity of the parameter space explored
by the Monte Carlo solvers achieves better fine-tuning of the
action policies and estimation of expected reward outcomes.
That is, the purposeful stochasticity of policy and value estimation
in MDPs provides a candidate explanation for why humans
have evolved imperfect noisy memories
as the more advantageous adaptation.
In sum, episodic scene construction according to the semantic
account is lacking an explicit time and incentive model,
both of which are integral parts of MDP frameworks.

\subsection{Sentinel hypothesis}
The DMN regions have been associated with processing the
experienced or expected relevance of environment stimuli
\citep{montague2006imaging}.
Processing self-relevant information was perhaps the first
cognitive account that was proposed for DMN function
\citep{gusnard2001medial, raichle2001pnas}.
Since then,
many investigators have speculated that neural activity in the DMN
may reflect the brain's continuous tracking of
relevance in the environment, such as spotting predators,
as an advantageous evolutionary adaptation \citep{randy2008, hahn2007cingulate}.
According to this cognitive account, the human brain's baseline maintains
a ``radar'' function to
detect subjectively relevant cues and unexpected events in the environment.
Propositions of a sentinel function to underlie DMN activity
have however seldom detailed
the mechanisms of
how attention and memory resources are exactly reallocated when
encountering a self-relevant environmental stimulus.
In the present MDP account,
promising action trajectories
are recursively explored by the human DMN. Conversely,
certain branches of candidate action trajectories
are detected to be less worthy to become mentally explored.
This mechanism, expressed by the Bellman equation,
directly implies stratified allocation of attention and working memory load
over relevant cues and events in the environment.
%
Further,
our account provides a parsimonious explanation for
the consistently observed DMN implication
in certain goal-directed experimental tasks and
in task-unconstrained mind-wandering \citep{smith2009, bzdok2016formal}.
Both environment-detached and environment-engaged
cognitive processes may entail DMN recruitment
if real or imagined experience is processed, manipulated, and
used for predictions.
During tasks,
the policy and value estimates may be
updated to optimize especially short-term action.
At rest, these parameter updates may
improve especially mid- and long-term action.
This horizon of the agent
is expressed in the $\gamma$ parameter in the MDP account.
We thus provide answers for the currently unsettled question why the involvement
of the same neurobiological brain circuit (i.e., DMN) has been documented
for specific task performances and baseline house-keeping functions.

More specifically,
environmental stimuli especially important for humans are frequently of
social nature. This is probably unsurprising
given that
the complexity of the social systems
is a likely to be a human-defining property
\citep{tomasello2009cultural}.
According to the ``social brain hypothesis'',
the human brain has especially been shaped for
forming and maintaining increasingly complex
social systems,
which allows solving ecological problems by social relationships
\citep{whiten1988machiavellian}.
Indeed, social topics amounted to roughly
two thirds of human everyday communication \citep{dunbar1997human},
while
mind-wandering at daytime and dreams during sleep
are rich in stories about people and
the complex relationships between them.
%
In line with this, the DMN has been proposed to be specialzed in
continuous processing of social information as a
physiological baseline of human brain function
\citep{schilbach2008minds}. This view was later challenged by observing
analogues of the DMN in monkeys \citep{mantini2011default},
cats \citep{popa2009contrasting},
and rats \citep{lu2012rat}, three species with
supposedly less advanced social-cognitive capacities.
Rather than functional specificity for processing social information,
the present formal account can parsimoniously incorporate
the dominance of social content in
human mental activity
as high value function estimates for information about humans
\citep{baker2009action, kampe2001psychology, krienen2010clan}.
The DMN may thus modulate reward processing
in the human agent in a way that prioritizes
appraisal of and action towards social contexts,
without excluding relevance of environmental cues of the physical world.
In sum,
our account on the DMN directly implies
its previously proposed ``sentinel'' function
of monitoring the environment for self-relevant information
in general and
inherently accommodates the importance of social environmental cues
as a special case.

\subsection{The free-energy principle and active inference}
% \subsubsection{The theory}
In Friston's free-energy principle (FEP) \citep{friston2010free,fristonAIorRL}, the brain is portrayed
as biomechanical inference engine which must minimize the long-term average of surprise.
Precursors of this theory can be traced back to \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial \textit{directional deep belief-net (DBN)}. According to FEP's account, the goal of the brain is then to optimize over a generative model $G$ of sensations: to iteratively
modify its internal representation $p_G(\z|\x)$ about objects in the world, their interactions and dynamics, etc., so as to minimize surprise when these representations are confronted with sensory input $\x$ during perception cycles. These is called the \textit{generative} model. FEP also postulates a dual model called the \textit{recognition} model,
which works in tandem with the generative model $p_R(\z|\x)$, to accomplish approximate inference. The recognition model dreams imaginary worlds $\z$,
while the latter tries to generate sensations $\x$ which match these dreams!

@@

% \paragraph{Comparison to our proposed theory.}
On the surface, a common point between the FEP and our proposed RL-based framework
  is placing the minimization of a surprise signal at the core of brain function.
  Indeed in RL, surprise minimization is subsumed  by accurate prediction of
  rewarding outcomes in the future. A ``free-energy'' agent is barely a biomechanical machine which has the tendency to resist undesired / harmful phase-transitions.
  \citep{friston2010free,fristonAIorRL,ortega2013thermodynamics}. Such a theory cannot by itself,
  explain the emergence of strategic behavior inherent in humans (e.g.,  cite dark-room experiment).

>>>>

Normal and disturbed metabolic fluctuations in the
human PMC have been closely related to
changes of conscious awareness \citep{cavanna2006precuneus}.
Indeed,
the PMC matures relatively late (i.e., myelination) during postnatal development in monkeys
\citep{goldman1987development}, which is generally considered to
be a sign of evolutionary sophistication \citep{flechsig1920}.
%
This DMN node has long been speculated to
reflect constant computation of
environmental statistics and its internal representation
as an inner ``mind's eye'' \citep{cavanna2006precuneus, leech_pcc2014}.
For instance, B\'alint's syndrome is a neurological disorder of conscious
awareness that results from medial damage in the parietal cortex
\citep{balint1909seelenlahmung}.
Patients are plagued by an
inability to combine various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as an inability to direct action towards
currently unattended environmental objects
(i.e., optic ataxia).
This can be viewed as a high-level impairment in gathering
information about alternative objects (i.e., exploration) as well as
leveraging these environmental opportunities (i.e., exploitation).
Congruently,
the human PMC was coupled in two functional connectivity analyses
\citep{bzdok2015subspecialization}
with the amygdala, involved in significance evaluation, and
the nucleus accumbens (NAc), involved in reward evaluation.
Specifically, among all parts of the PMC,
the ventral posterior cingulate cortex was
most connected to the laterobasal
nuclei group of the amygdala
\citep{bzdok2015subspecialization}.
This amygdalar subregion has been proposed to
continuously scan environmental input
for biological significance assessment
\citep{amygdala_bzdok}.

@@

\section{Relation to existing cognitive accounts on default mode function}
\subsection{Predictive coding hypothesis}
Predictive coding mechanisms
\citep{clark2013whatever, friston2008hierarchical}
are a frequently evoked idea in the context of default mode function
\citep{bar2007}.
Cortical responses are explained as
emerging from continuous functional interaction between
higher and lower levels of the neural processing hierarchy.
Feed-forward sensory processing is constantly calibrated by
top-down modulation from more multi-sensory and associative brain regions
further away from primary sensory cortical regions.
The dynamic interplay between cortical processing levels
may enable inferences about the world by reconciling
gaps between fresh sensory input and expectations computed
based on stored prior information.
At each stage of neural processing,
an internally generated prediction of aspects of environmental sensations is
directly compared against the actual environmental input.
A prediction error at one of the processing levels
incurs plasticity changes of neuronal projections
(i.e., adapting model parameters)
to allow for gradually improved future prediction of the environment.
In this way,
the predictive coding hypothesis offers explanations for
the constructive, non-deterministic nature of sensory perception
\citep{friston2010free, buzsaki2006rhythms} and
the intimate relation of motor action to sensory expectations
\citep{wolpert1995internal, kording2004bayesian}.
Contextual integration of sensorimotor perception-action cycles
may be maintained by top-down modulation
using a-priori information about the environment.

In short, predictive coding processes
conceptualize updates of the internal representation of the environment
to best accommodate and prepare the organism
for processing the constant influx of sensory stimulation and
performing action on the environment.
There are hence a number of common properties
between the predictive coding account
and the proposed formal account based on MDPs.
Importantly,
a generative model of how perceived sensory cues arise in the
world would be incorporated into
the current neuronal wiring.
Further,
both functional accounts are plausibilized by
neuroscientific evidence that suggest
the human brain to be a ``statistical organ'' \citep{friston2014phantastic}
with the biological purpose to
generalize from the past to new experiences.
Neuroanatomically, axonal
back projections indeed outnumber by far the axonal input projections
existing in the monkey and probably also human brain
\citep{salin1995corticocortical}.
These many and diverse modulatory influences from higher cortical areas
can inject prior knowledge
at every stage of processing environmental information.
Moreover,
both accounts provide a parsimonious explanation why the
human brain decreases processing load devoted to incoming information
when the environment is predictable.
This is because the internal generative
model only requires updates after descrepancies have occurred between
environmental reality and its internally reinstantiated representation.
Increased computation resources are however allocated
when unknown stimuli or
unexpected events are encountered by the organism.
The predictive coding and MDP account
naturally evoke a mechanism of brain plasticity in that
neuronal wiring gets increasingly adapted
when faced by unanticipated environmental challenges.

While sensory experience is a constructive process from both views,
the predictive coding account frames
sensory perception of the external world
as a generative experience due to the modulatory top-down influence at
various stages of sensory input processing.
This top-down generative design is replaced in our MDP view of the DMN
by a sequential decision-making framework.
Further,
the hierarchical processing aspect from predictive coding
is re-expressed in our account in form of
nested prediction of probable upcoming actions, states, and rewards.
While both accounts capture the consequences of action,
the predictive coding account is typically explained without
explicit parameterization of the agent's horizon and
has a tendency to be presented as emphasizing prediction about the
immediate future.
In the present account, the horizon of that
look into the future is made explicit in the $\gamma$ parameter
of the Bellman equation.
% \item \suggestremove{Mismatch negativity of PC is immanent in reward prediction error
% by the difference between actually predicted reward of an action given
% a state and the reward predicted by non-linear regression with graded
% discounting of future rewards.}
Finally,
the process of adapting the neuronal connections
for improved top-down modulation
takes the concrete form of stochastic gradient computation and
back-propagation in our MDP implementation.
It is however important to note that
the neurobiological plausibility of
the back-propagation procedure is controversal
\citep{goodfellow2016deep}.

In sum,
the present MDP account of DMN function may serve as a concrete implementation of
the long-standing predictive coding account from cognitive neuroscience.
MDPs have the advantage of exposing an explicit
mechanism for the horizon of future considerations and for
how the internal representation of the world is updated,
as well as
more explicitly reconciles sensory input processing and action considerations.
%
Recasting DMN function in terms of MDPs therefore naturally incorporates
a majority of aspects from the prediction coding hypothesis.

\subsection{Semantic hypothesis}
Another frequently proposed cognitive account to explain DMN function revolves
around forming logical associations and abstract analogies between
experiences and conceptual knowledge derived from past behavior
\citep{bar2007proactive, binder1999conceptual, constantinescu2016organizing}.
Analogies might naturally tie incoming new sensory stimuli to
explicit world knowledge (i.e., semantics)
\citep{bar2009proactive}.
The encoding of complex environmental features could thus be facilitated
by association to similar states.
%
Going beyond isolated meaning and concepts extracted from the world,
semantic building blocks may need to get recombined to enable
mentally envisioning non-existing scenarios.
As such, semantic knowledge would be a prerequisite for optimizing behavior
by constantly simulating possible future scenarios
\citep{boyer2008evolutionary, binder2011neurobiology}.
Such cognitive processes can afford
the internal construction and elaboration of necessary information
that is not presented in the surrounding sensory environment
by recombining building blocks of
concept knowledge and episodic memories
\citep{hassabis2009construction}.
Indeed, in aging humans, remembering the past and imaging the future
equally decreased in the level of detail and were associated with
concurrent deficits in forming and integrating relationships between
items \citep{addis2008age, spreng2006temporal}.
Further,
% a constructive account can explain the reciprocal relationship
% between an egocentric first person perspective and
% an allocentric bird’s eye perspective immersed in
% self-reflection, semantic associations, and autobiographical memories.
% %
% Cognitive aspects of egocentric-allocentric switching
episodic memory, language, problem solving,
planning, estimating other people's thoughts, and spatial navigation
represent neural processes that are likely to
build on abstract world knowledge and logical associations
for integrating the constituent elements in rich and coherent mental scenes
\citep{schacter2007remembering}.
Such scene construction processes could contribute to interpreting the
present and foretelling the future.
Further,
mental scene construction has been proposed
to imply a distinction between
engagement in the sensory environment
and internally generated mind-wandering
\citep{buckner2007self}.
These investigators stated that
``A computational model [...] will probably require a form of
regulation by which perception of the current world is suppressed
while simulation of possible alternatives are constructed,
followed by a return to perception of the present.''.
The semantic hypothesis is for instance supported by evidence in animals that
could learn a \textit{cognitive map of the environment}
and exploit it later for other means \citep{tolman1948cognitive}.

In comparison,
both the semantic hypothesis and the present formal account based on MDPs
expose mechanisms of how action considerations could be mentally explored.
In both accounts,
there is also no reason to assume that predictions of various
levels of complexity, abstraction, timescale, and purpose
rely on mechanisms that are qualitatively different. This concurs with
DMN activity increases across time, space, and content domains
demonstrated in many neuroimaging studies
\citep{spreng2009common, laird2009, bzdok2012morality, binder2009}.
Further, the semantic hypothesis
and MDP account provide explanations why HC damage does
not only impair recalling memories, but also hypothetical and future
thinking \citep{hassabis2007patients}.
While both semantic hypothesis and
our formal account propose memory-based internally
generated information for probabilistic mental models of action outcomes,
MDPs render explicit the grounds on which an action is
eventually chosen, namely, the estimated cumulative reward.
In contrast to many versions of the semantic hypothesis,
the MDPs naturally integrate the egocentric view
(more related to current action, state, and reward) and the
world view (more related to past and future actions, states, and rewards)
on the world in a same optimization problem.
Finally,
the semantic account of DMN function does not offer
a mechanistic explanation \textit{how}
explicit world knowledge and semantic analogies thereof
lead to prediction of future actions and states.
The semantic hypothesis does not explain why memory recall
for scene construction in humans is typically fragmentary and noisy
instead of accurate and reliable.
In contrast to existing accounts on semantics and
mental scene construction, the random and creative aspects of DMN function
are explained in MDPs by the advantages of stochastic optimization.
Our MDP account provides an algorithmic explanation in that
stochasticity of the parameter space explored
by Monte Carlo approximation achieves better fine-tuning of the
action policies and estimation of expected reward outcomes.
That is, the purposeful stochasticity of policy and value estimation
in MDPs provides a candidate explanation for why humans
have evolved imperfect noisy memories
as the more advantageous adaptation.
In sum, episodic scene construction according to the semantic
account is lacking an explicit time and incentive model,
both of which are integral parts of MDP frameworks.

\subsection{Sentinel hypothesis}
The DMN regions have been associated with processing the
experienced or expected relevance of environment cues
\citep{montague2006imaging}.
Processing self-relevant information was perhaps the first
cognitive account that was proposed for DMN function
\citep{gusnard2001medial, raichle2001pnas}.
Since then,
many investigators have speculated that neural activity in the DMN
may reflect the brain's continuous tracking of
relevance in the environment, such as spotting predators,
as an advantageous evolutionary adaptation \citep{randy2008, hahn2007cingulate}.
According to this cognitive account, the human brain's baseline maintains
a ``radar'' function to
detect subjectively relevant cues and unexpected events in the environment.
Propositions of a sentinel function to underlie DMN activity
have however seldom detailed
the mechanisms of
how attention and memory resources are exactly reallocated when
encountering a self-relevant environmental stimulus.
In the present MDP account,
promising action trajectories
are recursively explored by the human DMN. Conversely,
certain branches of candidate action trajectories
are detected to be less worthy to become mentally explored.
This mechanism, expressed by the Bellman equation,
directly implies stratified allocation of attention and working memory load
over relevant cues and events in the environment.
%
Further,
our account provides a parsimonious explanation for
the consistently observed DMN implication
in certain goal-directed experimental tasks and
in task-unconstrained mind-wandering \citep{smith2009, bzdok2016formal}.
Both environment-detached and environment-engaged
cognitive processes may entail DMN recruitment
if real or imagined experience is processed, manipulated, and
used for predictions.
During tasks,
the policy and value estimates may be
updated to optimize especially short-term action.
At rest, these parameter updates may
improve especially mid- and long-term action.
This horizon of the agent
is expressed in the $\gamma$ parameter in the MDP account.
We thus provide answers for the currently unsettled question why the involvement
of the same neurobiological brain circuit (i.e., DMN) has been documented
for specific task performances and baseline house-keeping functions.

More specifically,
environmental stimuli especially important for humans are frequently of
social nature. This is probably unsurprising
given that
the complexity of the social systems
is a likely to be a human-defining property
\citep{tomasello2009cultural}.
According to the ``social brain hypothesis'',
the human brain has especially been shaped for
forming and maintaining increasingly complex
social systems,
which allows solving ecological problems by social relationships
\citep{whiten1988machiavellian}.
Indeed, social topics amounted to roughly
two thirds of human everyday communication \citep{dunbar1997human},
while
mind-wandering at daytime and dreams during sleep
are rich in stories about people and
the complex relationships between them.
%
In line with this, the DMN has been proposed to be specialized in
continuous processing of social information as a
physiological baseline of human brain function
\citep{schilbach2008minds}. This view was later challenged by observing
analogues of the DMN in monkeys \citep{mantini2011default},
cats \citep{popa2009contrasting},
and rats \citep{lu2012rat}, three species with
social-cognitive capacities that are supposedly less advanced than in humans.
Rather than functional specificity for processing social information,
the present MDP account can parsimoniously incorporate
the dominance of social content in
human mental activity
as high value function estimates for information about humans
\citep{baker2009action, kampe2001psychology, krienen2010clan}.
The DMN may thus modulate reward processing
in the human agent in a way that prioritizes
appraisal of and action towards social contexts,
without excluding relevance of environmental cues of the physical world.
In sum,
our account on the DMN directly implies
its previously proposed ``sentinel'' function
of monitoring the environment for self-relevant information
in general and
inherently accommodates the importance of social environmental cues
as a special case.

\subsection{The free-energy principle and active inference}
% \subsubsection{The theory}
In Friston's free-energy principle (FEP) \citep{friston2010free,fristonAIorRL}, the brain is portrayed
as biomechanical inference engine
dedicated to minimizing the long-term average of surprise.
Precursors of this theory can be traced back to \citep{dayan1995helmholtz} in which they introduced the so-called \textit{Helmholtz machine}, a hierarchical factorial \textit{directional deep belief-net (DBN)}. According to the FEP account,
the goal of the brain is then to optimize over a generative model $G$ of sensations: to iteratively
modify its internal representation $p_G(\z|\x)$ about objects in the world, their interactions and dynamics,
so as to minimize surprise when these representations are confronted with sensory input $\x$ during perception cycles. These is called the \textit{generative} model. FEP also postulates a dual model called the \textit{recognition} model,
which works in tandem with the generative model $p_R(\z|\x)$, to accomplish approximate inference. The recognition model dreams imaginary worlds $\z$,
while the latter tries to generate sensations $\x$ which match these dreams!

@@

% \paragraph{Comparison to our proposed theory.}
On the surface, a common point between the FEP and our proposed RL-based framework
  is placing the minimization of a surprise signal at the core of brain function.
  Indeed in RL, surprise minimization is subsumed  by accurate prediction of
  rewarding outcomes in the future. A ``free-energy'' agent is barely a biomechanical machine which has the tendency to resist undesired phase-transitions
  \citep{friston2010free,fristonAIorRL,ortega2013thermodynamics}. Such a theory cannot by itself,
  explain the emergence of strategic behavior inherent in humans (e.g.,  cite dark-room experiment).
